---
title: "Data Visualisation"
author: "DCS"
format: 
  revealjs:
    chalkboard: TRUE
    incremental: true
    logo: cdcs_logo.JPG
    footer: "[www.cdcs.co.uk](www.cdcs.co.uk)"
editor: visual
---


```{r setup, echo = FALSE}

#install.packages(c("extrafont","plotly","gghalves" ))

library(extrafont) #additional fonts - will not work in noteable, but useful for own device! 
# font_import() #gain access to additional fonts - WARNING - will take a long time
#loadfonts(device = "win") #load fonts into R - Mac users: change "win" to "all".

library(tidyverse) # data tidying
library(plotly) # interactive plot
library(gghalves) # half plots - useful for distributions



Whisky_data <- read_csv(url("https://raw.github.com/DCS-training/SummerSchoolStream2ContentTesting/main/Day1/RMD_file_data/tidied_whisky_data.csv")) %>%
  mutate_if(is.character, as.factor)

```

## Grammar of Graphics {#sec-grammar-of-graphics}

-   A set of rules to facilitate data visualisation developed by [Leland Wilkinson](https://link.springer.com/book/10.1007/0-387-28695-0).
-   It allows us to think of plot generation by following a step by step recipe (and like all good recipes, can be modified when needed).
-   For a full overview, check out [Hadley Wickham's free textbook](https://ggplot2-book.org/layers#introduction) on using ggplot.
-   For inspiration on more complex plots and advanced techniques, [Cedric Sherer's blog](https://www.cedricscherer.com/tags/ggplot2/) is full of great ideas!

## Grammar of Graphics

The recipe of a nice tasty plot includes:

1)  The data
2)  Aesthetic Mapping
3)  Geometric Objects and Statistical transformations
4)  Scaling and Faceting
5)  Themes

## What does this mean? {#sec-what-does-this-mean}

-   [GGplot (grammar of graphics plot)](https://www.tandfonline.com/doi/abs/10.1198/jcgs.2009.07098) has the grammar of graphics built into its structure.
-   This allows us to build plots in layers.
-   The plots start from the data, and build all the way up to the decorations.
-   This provides us with a lot of control to get the plots exactly as we want them.

## What does this mean?

```{r fancyplot, echo = FALSE}
#| fig-width: 16
#| fig-height: 9
fancy_plot <-  Whisky_data %>% # assigning our code to "plot object"
  filter(category  == "Single Malt Scotch")%>% 
  count(Owner) %>%
  top_n(n = 2) %>% 
  left_join(Whisky_data) %>%
ggplot(
  aes(x = Region, y = review.point, fill = Region)) + 
 # geom_violin() +
  geom_boxplot()+
  stat_summary(fun.data = mean_se,
    color = "black", linewidth = 1, size = 1)+
  stat_summary(fun.data = mean_se,
    color = "white", linewidth = .7, size = .7)+
  facet_wrap(~reorder(Owner, -n), 
    ncol = 1, 
    scales = "free_y", 
    strip.position = "top") +#
  coord_flip()+
  scale_fill_manual(
    values = c(Speyside = "#6072A0", Lowland = "#C7AC59", Highland = "#6F3A7E", 
               Islay = "#A84046", Island = "#D40A9F", Campbeltown = "#3A8259" )) + 
  theme_bw() + 
  theme(strip.background = 
    element_rect(fill="#EAF3FD",
    color = "black",
    linewidth = .7),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.background = element_rect(fill = '#F9F9F9', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "top",
    text= element_text( # command to control text
    size=12, #font size
    family="Franklin Gothic Medium", #font style
    face = "bold")) + # choose bold/italics/default
  labs(x = "",
    y = "Online Review Scores",
    title = "Whisky Review Scores by Region from the Top 2 largest Producers")
  fancy_plot
ggplotly(fancy_plot) 

```

## Building our Plot

-   Our first call is to use the function `ggplot()`
-   By itself, it looks like this:

```{r ggplot_function, echo = TRUE}
#| output-location: column-fragment
ggplot()
```

## Data layer

-   From there, we need to specify the data we need.
-   We can feed in the data as it is.
-   Or we can wrangle the data specifically for the visualisation.
-   However, at this stage our plot is still blank.

## Data Layer

```{r data_layer, echo = TRUE}
#| code-line-numbers: "2"
#| output-location: column-fragment
#| fig-width: 10
#| fig-height: 6

# Regular data
ggplot(data = Whisky_data)


```

## Adding our coordinates

-   Our plots are empty as we've not told *R* what variables to use.
-   Thankfully, the command is very simple, using the `aes()` command after specifying the data.
-   Here we simply label our *x* and *y* coordinates.
-   But as we will see, we still don't have any visual ques in the plot itself.

## Adding our coordinates

```{r coordinates, echo = TRUE}
#| code-line-numbers: "3"
#| output-location: column-fragment
#| fig-width: 10
#| fig-height: 6

# Standard version
ggplot(data = Whisky_data,
  aes(x = Region, y = review.point)) # setting x and y

```

## Graphic element of the plot

-   Here we link the **coordinates** from our **data** to visual geometries.
-   We use geoms (geometric functions) to decide how to shape the coordinates.
-   This allows us to shape our plot with our determined coordinates.
-   We can also apply statistical functions with the geoms.
-   We can place multiple geom layers.
-   The order of our layers is determined by the order we code.

## Graphic element of the plot {#sec-graphic-element-of-the-plot}

```{r geoms, echo = TRUE}
#| code-line-numbers: "4"
#| output-location: column-fragment
#| fig-width: 8
#| fig-height: 4

# First Layer
ggplot(data = Whisky_data,
  aes(x = Region, y = review.point)) + # setting x and y 
  geom_half_boxplot()
```

```{r geoms_2, echo = TRUE}
#| code-line-numbers: "5"
#| output-location: column-fragment
#| fig-width: 10
#| fig-height: 6
# Second Layer
ggplot(data = Whisky_data,
  aes(x = Region, y = review.point))+
  geom_half_boxplot(size = .5, width = .2)+ # scaling size and width
  geom_half_violin(side = "r") #choosing right side for aesthetics


```

## Adding statistical layers

-   Here we can model visualise any model we've used in our analyses.
-   This is a vital step for communicating our research.
-   It's also a fundamental step in validating our findings.
-   R has so many options for visualising our models! From loess and linear models, to means and standard deviations.

## Adding statistical layers

```{r geoms_3, echo = TRUE}
#| code-line-numbers: "6|8"
#| output-location: column-fragment
#| fig-width: 10
#| fig-height: 6
# Second Layer
ggplot(data = Whisky_data,
  aes(x = Region, y = review.point))+
  geom_half_boxplot(size = .5, width = .2)+ # scaling size and width
  geom_half_violin(side = "r") + #choosing right side for aesthetics
  stat_summary(fun.data = mean_se,
    color = "black", linewidth = 1.2, size = 1.2)+ # adjusting size
  stat_summary(fun.data = mean_se, 
    color = "white", linewidth = .7, size = .7) # making smaller for a white centre

```

## Scaling our plots

-   This is closer, but we can see that our plots need some serious work.
-   We have an influential outlier that's causing havoc.
-   We can also account for other influential variables (like like owners).
-   We can focus our data for a clearer narrative.
-   We can also tweak the coordinates.

## Scaling our plots

```{r scaled_geoms, echo = TRUE}
#| code-line-numbers: "2|3|4|5"
#| output-location: column-fragment
#| fig-width: 12
#| fig-height: 6

Whisky_data %>% # 
  filter(category  == "Single Malt Scotch") %>% #filter to include only single malt
  count(Owner) %>% # counting Owner
  top_n(n = 2) %>% # so we get top 2 Owners by count
  left_join(Whisky_data) %>%  # joining back to data
ggplot(
  aes(x = Region, y = review.point))+
  geom_half_boxplot(size = .5, width = .2)+ 
  geom_half_violin(side = "r") +
  stat_summary(fun.data = mean_se, 
    color = "black", linewidth = 1.2, size = 1.2)+
  stat_summary(fun.data = mean_se,
    color = "white", linewidth = .7, size = .7)
  
```

## Scaling our plots

```{r scaled_geoms_2, echo = TRUE}
#| code-line-numbers: "8|15|16|17|18|19"
#| output-location: column-fragment
#| fig-width: 12
#| fig-height: 10

Whisky_data %>% # 
  filter(category  == "Single Malt Scotch") %>% 
  count(Owner) %>% 
  top_n(n = 2) %>% 
  left_join(Whisky_data) %>% 
ggplot(
  aes(x = Region, y = review.point, 
      fill = Region))+
  geom_half_boxplot(size = .5, width = .2)+ 
  geom_half_violin(side = "r") +
  stat_summary(fun.data = mean_se, 
    color = "black", linewidth = 1, size = 1)+
  stat_summary(fun.data = mean_se, 
    color = "white", linewidth = .6, size = .6) +
  facet_wrap(~reorder(Owner, -n), # faceting by Owner, ranking by descending n
    ncol = 1, # number of columns 
    scales = "free_y", #changed to "free_y" to drop unused labels
    strip.position = "bottom") + #label position +
  coord_flip() # flipping chart on its side
```

## Setting themes and color schemes

-   Some journals will require certain stylistic designs for your figures.
-   This will include colour schemes, but also includes fonts and other design features.
-   Thankfully ggplot allows us to control these aspects of the plot with full customisation options as well.
-   Here we will inspire our colour scheme from The Scotch Whisky Experience tour material colour palate (and I use Power Point to get the colour hex codes).

## Setting themes and color schemes

-   Pre-made colour and theme options also exist:
-   [Journal color schemes](https://cran.r-project.org/web/packages/ggsci/vignettes/ggsci.html#Introduction)
-   [Theme options](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)
-   [Wes Anderson Color Scheme](https://github.com/karthik/wesanderson)

## Setting themes and color schemes

### Decluttering our code

```{r taming_code, echo = TRUE}
#| code-line-numbers: "1|20"
#| output-location: column-fragment
#| fig-width: 10
#| fig-height: 10
plot <- Whisky_data %>% 
  filter(category  == "Single Malt Scotch") %>% 
  count(Owner) %>% 
  top_n(n = 2) %>% 
  left_join(Whisky_data) %>% 
ggplot(
  aes(x = Region, y = review.point, fill = Region))+
  geom_half_boxplot(size = .5, width = .2)+ 
  geom_half_violin(side = "r") +
  stat_summary(fun.data = mean_se, 
    color = "black", linewidth = 1, size = 1)+
  stat_summary(fun.data = mean_se,
    color = "white", linewidth = .6, size = .6) +
  facet_wrap(~reorder(Owner, -n), 
    ncol = 1, 
    scales = "free_y", 
    strip.position = "bottom") + 
  coord_flip()

plot
```

## Setting themes and color schemes

```{r themes_and_color_schemes, echo = TRUE}
#| code-line-numbers: "2|3|4|5|6|7|8|9|10|11|12"
#| output-location: column-fragment
#| fig-width: 10
#| fig-height: 10

plot +
  scale_fill_manual(
    values = c(Speyside = "#6072A0", Lowland = "#C7AC59", Highland = "#6F3A7E", 
               Islay = "#A84046", Island = "#D40A9F", Campbeltown = "#3A8259" )) + #setting to discrete variables
  theme_bw() + # simpler theme
  theme(strip.background = # play with facet label
    element_rect(fill="#EAF3FD",# facet label fill
    color = "black"), #facet label border
    legend.position = "bottom", # change legend position,
    panel.background = element_rect(fill = '#F9F9F9', colour = 'black'), # background and border
    panel.grid.major = element_blank(), # removing gridlines
    panel.grid.minor = element_blank()  # removing gridlines
        ) 
```

## Setting themes and color schemes - font

```{r, font, echo = TRUE}
#initialising fonts - 
## Uncomment and copy + paste, and run this step in console
## Its too slow and intensive to run through quatro/markdown itself
# library(extrafont) # library for fonts tools
# font_import() #gain access to additional fonts
# loadfonts(device = "win") #load fonts into R
```

## Setting themes and color schemes: Font

```{r themes_and_color_schemes_font, echo = TRUE}
#| code-line-numbers: "9|13|14|15"
#| output-location: column-fragment
#| fig-width: 10
#| fig-height: 10

better_plot <- plot +
  scale_fill_manual(
    values = c(Speyside = "#6072A0", Lowland = "#C7AC59", Highland = "#6F3A7E", 
               Islay = "#A84046", Island = "#D40A9F", Campbeltown = "#3A8259" )) + 
  theme_bw() + 
  theme(strip.background = 
    element_rect(fill="#EAF3FD",
    color = "black"),
    legend.position = "bottom",
    panel.background = element_rect(fill = '#F9F9F9', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    text= element_text( # command to control text
    size=16, #font size
    family="Comic Sans MS") #font style
        ) 

better_plot
```

## Setting themes and color schemes - font

```{r lFont_change, echo = TRUE}
#| code-line-numbers: "15|16"
#| output-location: column-fragment
#| fig-width: 10
#| fig-height: 10

better_plot <- plot +
  scale_fill_manual(
    values = c(Speyside = "#6072A0", Lowland = "#C7AC59", Highland = "#6F3A7E", 
               Islay = "#A84046", Island = "#D40A9F", Campbeltown = "#3A8259" )) + 
  theme_bw() + 
  theme(strip.background = 
    element_rect(fill="#EAF3FD",
    color = "black"),
    legend.position = "none", # dropping legend, as label is clear
    panel.background = element_rect(fill = '#F9F9F9', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    text= element_text( # command to control text
    size=16, #font size
    family="Franklin Gothic Medium", #font style
    face = "bold"), # choose bold/italics/default
        ) 

better_plot
```

## Labelling our plot

-   Our plot is now almost ready to go!
-   We just need to customise our labels.
-   This helps make our plots more informative

## Labelling our plot

```{r labelled_plot, echo = TRUE}
#| code-line-numbers: "3|4|5|6|7"
#| output-location: column-fragment
#| fig-width: 10
#| fig-height: 10

fancy_plot<- better_plot +
  labs(
       x = "",
       y = "Review Scores",
       title = "Whisky Review Scores by Region from the Top 2 largest Producers")

fancy_plot
```

## Labelling our plot

```{r}
#| fig-width: 16
#| fig-height: 10
fancy_plot
```

## Bonus - Interactive plot

-   Using the `plotly` package, we can make our plots interactive when in html format.
-   This can be very useful for presenting our data to a live audience.
-   It can also be useful for teaching purposes - to help students understand what our coefficients mean visually.
-   Note that some features of the plot maybe lost, as `plotly` uses a slightly different structure to ggplot.
-   Otherwise, it is very easy to convert our plot with the `ggplotly()` function

## Bonus - Interactive plot

```{r interactive_plot, echo = TRUE}
#| output-location: column-fragment
#| fig-width: 12
#| fig-height: 10

p <-  Whisky_data %>% # assigning our code to "plot object"
  filter(category  == "Single Malt Scotch")%>% 
  count(Owner) %>%
  top_n(n = 2) %>% 
  left_join(Whisky_data) %>%
  ggplot(
    aes(x = Region, y = review.point, fill = Region))+
      geom_boxplot()+ 
  scale_fill_manual(
    values = c(Speyside = "#6072A0", Lowland = "#C7AC59", Highland = "#6F3A7E", 
               Islay = "#A84046", Island = "#D40A9F", Campbeltown = "#3A8259" )) +
    stat_summary(fun.data = mean_se, 
      color = "black", linewidth = 1, size = 1)+
    stat_summary(fun.data = mean_se, 
      color = "white", linewidth = .6, size = .6) +
  facet_wrap(~reorder(Owner, -n) , 
      ncol = 1, 
      scales = "free_y", 
      strip.position = "bottom") + 
  coord_flip() + 
  theme_bw()+
  theme(strip.background = 
    element_rect(fill="#EAF3FD",
    color = "black",
    linewidth = .4),
    text= element_text( # command to control text
    size=16, #font size
    family="Franklin Gothic Medium", #font style
    face = "bold"), # choose bold/italics/default
    panel.background = element_rect(fill = '#F9F9F9', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
        ) + 
  labs(x ="",
       y = "Online Review Scores",
       title = "Whisky Review Scores by Region from the Top 2 largest Producers")

p # ggplotly(p)
```

## Bonus - Interactive plot

```{r}
#| output-location: slide
#| fig-width: 16
#| fig-height: 9

ggplotly(p)
# this is the code
# We had to comment it out, as plotly didnt play nicely.

```
